SU($n$) spin models near the Heisenberg critical point

### 2020-03-22 collective_sun/sun_perturbation.tex
rewrote derivation of diagrammatic expansion of operator products

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\P}{\mathcal{P}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\Z}{\mathcal{Z}}

@@

In order to induce non-trivial dynamics for initially PS states
$\ket\psi\in\M_0$, which are otherwise stationary eigenstates of the
interaction Hamiltonian $H_0$ in \eqref{eq:H_0}, we will add to the
Hamiltonian an operator of the general form \red{[TODO: consider more
  general sums of such operators]}
\begin{align}
  X_{\m w} \equiv \sum_{k\in\C_N\p{M}} w_k X_k,
  \label{eq:pert}
\end{align}
where $X$ is an $M$-spin operator that is invariant under arbitrary
permutation of its tensor factors; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
a choice of $M$ distinct spins that $X_k$ acts on; $\C_N\p{M}$ is the
set of all choices of $M$ elements from a set of size $N$; and $\m w$
is a dimension-$M$ tensor with scalar entries $w_k$.  Even though only
${N\choose M}$ entries of $\m w$ are used to define $X_{\m w}$ in
\eqref{eq:pert}, for later convenience we define $w_k$ for all indices
$k\in\ZZ_N^M$ by enforcing that $\m w$ is symmetric with respect to
any permutation of its indices, and that $w_k=0$ for any $k$ with
repeated values. \red{[TODO: relax the restrictions on $X$ and $\m w$
  to allow for more general perturbations, e.g.~of the form
  $s_\x\otimes s_\y-s_\y\otimes s_\x$]}

@@

Defining $\m w_\Delta$ as the projection of the tensor $\m w$ onto the
$\Delta$-eigenspace of $\hat{\m h}$, we can expand
\begin{align}
  \m w = \sum_\Delta \m w_\Delta,
  &&
  \hat{\m h}\p{\m w} = \sum_\Delta \Delta \m w_\Delta,
\end{align}
and in turn
\begin{align}
  X_{\m w} = \sum_\Delta X_{\m w_\Delta},
  &&
  X_{\hat{\m h}\p{\m w}} = \sum_\Delta \Delta X_{\m w_\Delta}.
\end{align}
Substituting these expansions into \eqref{eq:diagnosis}, we find
\begin{align}
  H_0 X_{\m w} \ket\psi
  = \sum_\Delta\p{E_0 + \Delta} X_{\m w_\Delta} \ket\psi,
\end{align}
which implies that the action of $X_{\m w_\Delta}$ on a PS state
$\ket\psi\in\M_0$ generates a state of definite interaction energy
$E_0+\Delta$.  It follows that
$\P_\Delta X_{\m w} \P_0 = \P_\Delta X_{\m w_\Delta} \P_0 = X_{\m
  w_\Delta} \P_0$, and so
\begin{align}
  H_{\t{eff}}^{(2)} = -\sum_{\Delta\ne0}
  \f1\Delta \P_0 X_{\m w_\Delta} X_{\m w_\Delta} \P_0.
\end{align}
The product $\P_0 X_{\m w_\Delta} X_{\m w_\Delta} \P_0$ generally
includes $r$-body operators for all positive integers $r\le2M$.  We
provide a general prescription for simplifying such products in
Appendix \ref{sec:operator_product}, and explicitly work out the
single-body ($M=1$) and two-body ($M=2$) cases in Appendices
\ref{sec:single_pair_prod} and \ref{sec:two_pair_prod}.  If $X$ is a
single-body operator ($M=1$), then
\begin{align}
  H_{\t{eff}}^{(2)}
  = \sum_{\Delta\ne0} \f{\var\p{\m w_\Delta}}{\p{N-1}\Delta}
  \p{\col{X}^2 - N \col{X^2}},
  &&
  \var\p{\m m} \equiv \f1N \sum_{p\in\ZZ_N} \p{m_p-\bar m}^2,
\end{align}
which in the SU(2) case implies that an inhomogeneous magnetic field
($X\propto\sigma_\z$) yields an OAT Hamiltonian at second order
perturbation theory.

@@

In order to project an operator $X_{\m w}$ onto
$\ul{\M}_{r_{\t{max}}}$, we first need to construct a basis for
$\ul{\M}_{r_{\t{max}}}$, and then compute all matrix elements of
$X_{\m w}$ in the constructed basis.  Our strategy will be to first
identify a basis for $\M_0$, and then find, for each
$r=1,\cdots,r_{\t{max}}$, a suitable set of $r$-body operators
$\set{O}$ and dimension-$r$ tensors $\set{\m u}$ for which the states
$\ket{O,\m u,m}\propto O_{\m u}\ket{m}$ form an orthonormal basis for
$\M_r$.  Matrix elements of $X_{\m w}$ with respect to this basis for
$\ul{\M}_{r_{\t{max}}}$ then take the form
\begin{align}
  \bk{O,\m u,\ell|X_{\m w}|Q,\m v,m}
  = \f{\bk{\ell|O_{\m u}^\dag X_{\m w} Q_{\m v}|m}}
  {\sqrt{\bk{\ell|O_{\m u}^\dag O_{\m u}|\ell}
      \bk{m|Q_{\m v}^\dag Q_{\m v}|m}}}.
\end{align}
In order to compute these matrix elements, we simplify products of the
form $\P_0 O_{\m u}^\dag X_{\m w} Q_{\m v} \P_0$ and
$\P_0 O_{\m u}^\dag O_{\m u} \P_0$, with $\P_0$ a projector onto
$\M_0$, in Appendix \ref{sec:operator_product}.

@@

@@

In this section, we simplify the products of the form
$H_0X_{\m w}\ket\psi$ to arrive at the expansion in
\eqref{eq:diagnosis}, which defines conditions under which a
multi-body perturbation $X_{\m w_\Delta}$ generates eigenstates of the
SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0} when
acting on a permutationally symmetric state $\ket\psi\in\M_0$.  Here
$X_{\m w}$ is an $M$-body perturbation defined by
\begin{align}
  X_{\m w} \equiv \sum_{k\in\C_N\p{M}} w_k X_k,
  \label{eq:pert_copy}
\end{align}
where $X$ is an $M$-spin operator that is invariant under arbitrary
permutation of its tensor factors; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
a choice of $M$ distinct spins that $X_k$ acts on; $\C_N\p{M}$ is the
set of all choices of $M$ elements from a set of size $N$; and $\m w$
is a dimension-$M$ tensor with scalar entries $w_k$.  Even though only
${N\choose M}$ entries of $\m w$ are used to define $X_{\m w}$ in
\eqref{eq:pert_copy}, for later convenience we define $w_k$ for all
indices $k\in\ZZ_N^M$ by enforcing that $\m w$ is symmetric with
respect to any permutation of its indices, and that $w_k=0$ for any
$k$ with repeated values. \red{[TODO: relax the restrictions on $X$
  and $\m w$ to allow for more general perturbations, e.g.~of the form
  $s_\x\otimes s_\y-s_\y\otimes s_\x$]}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Projecting products of multi-body operators onto the
  permutationally symmetric manifold}
\label{sec:operator_product}

@@

Given a system of $N$ spins, we wish to project a product of
multi-body operators onto the permutationally symmetric manifold
$\M_0$.  For $p$ multi-body operators, the projection of such a
product takes the form
\begin{align}
  \P_0 \sp{\prod_{j\in\ZZ_p}
    \sum_{k\in\C_N\p{M_j}} w_j\p{k} X_j\p{k}} \P_0,
  \label{eq:sym_prod_proj}
\end{align}
where $\P_0$ is a projector onto the PS manifold $\M_0$; $j$ indexes
one of the $p$ multi-body operators; $X_j$ is an $M_j$-spin operator
that is invariant under arbitrary permutation of its tensor factors;
$k\in\C_N\p{M_j}$ is a choice of $M_j$ spins that $X_j\p{k}$ acts on;
$\C_N\p{M_j}$ is the set of all choices of $M_j$ elements from a set
of size $N$; and $\m w_j$ are dimension-$M_j$ tensors with scalar
entries $w_j\p{k}$. \red{[TODO: relax the restrictions on $X_j$ and
  $\m w_j$ to allow for more general perturbations, e.g.~of the form
  $s_\x\otimes s_\y-s_\y\otimes s_\x$]}

To simplify the product in \eqref{eq:sym_prod_proj}, we first write
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} X_j\p{k}
  = \prod_{j\in\ZZ_p} \f1{M_j!} \sum_{k\in\ZZ_N^{M_j}}
  w_j\p{k} X_j\p{k},
  \label{eq:sym_prod_vec}
\end{align}
where we extend the definitions of $w_j\p{k}$ and $X_j\p{k}$ to all
vectors $k\in\ZZ_N^{M_j}$ by enforcing that these objects are
symmetric with respect to permutations of the entries in $k$, and that
$w_j\p{k}=X_j\p{k}=0$ if $k$ contains any repeated entries.  We then
collect terms whose operator content is equivalent up to a permutation
of spins.  To this end, we classify terms in \eqref{eq:sym_prod_vec}
by the numbers $g_S$ of indices shared by all tensors $w_j$ with
$j\in S\subset\ZZ_p$.  For example, a term of the form
$w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct indices
$\p{a,b,c,d,e}\in\C_N\p{5}$ would have
\begin{align}
  g_{\set{1}} = 1,
  &&
  g_{\set{1,2,3}} = 1,
  &&
  g_{\set{1,3}} = 1,
  &&
  g_{\set{2,3}} = 2,
\end{align}
and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
assignment can be represented by the Venn diagram
\begin{align}
  \diagram{example_123},
  \label{eq:venn_diagram}
\end{align}
where $g_S$ is determined the number of dots at the intersection of
circles $\set{w_j:j\in S}$.  Denoting the set of all nonempty subsets
of $\ZZ_p$ by $\PP^*\p{\ZZ_p}$, we denote an assignment of distinct
indices according to choices of $g_S$ for all $S\in\PP^*\p{\ZZ_p}$ by
$g$.  Given a vector $\m M$ of the dimensions $M_j$, we then denote
the set of all valid index assignments $g$ by $\G\p{\m M}$.  For
consistency, all valid index assignments $g\in\G\p{\m M}$ must assign
exactly $M_j$ indices to the dimension-$M_j$ tensor $w_j$, i.e.~
\begin{align}
  \sum_{S\in\PP^*\p{\ZZ_p}\,:\,j\in S} g_S
  = \sum_{R\in\PP^*\p{\ZZ_p\setminus\set{j}}} g_{\set{j}\cup R}
  = M_j
\end{align}
for all $j\in\ZZ_p$.  The set of valid index assignments $\G\p{\m M}$
is then essentially the set of all Venn diagrams of the form in
\eqref{eq:venn_diagram} with $p=\dim\p{\m M}$ circles and a total of
$M_j$ dots within circle $w_j$.

A classification of terms in \eqref{eq:sym_prod_vec} by index
assignments $g\in\G\p{\m M}$ allows us to expand
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\ZZ_N^{M_j}} w_j\p{k} X_j\p{k}
  \EQPS \sum_{g\in\G\p{\m M}} \S\p{g} w\p{g} X\p{g},
  \label{eq:sym_prod_group_start}
\end{align}
where $\EQPS$ denotes equality up to a restriction to the
permutationally symmetric manifold; $X\p{g}$ is an operator acquired
by assigning indices (choices of spins) to all $X_j$ in a manner
consistent with $g$; $w\p{g}$ is a scalar acquired by summing over all
indices assigned to the tensors $w_j$ according to $g$; and $\S\p{g}$
is a symmetry factor accounting for the number of ways to assign
indices according to $g$.

In order to write out the factors in \eqref{eq:sym_prod_group_start}
explicitly, we identify the set of values that the distinct indices
assigned according to $g$ can take:
\begin{align}
  \C_N\p{g}
  \equiv \set{ k \in \bigotimes_{S\in\PP^*\p{\ZZ_p}} \C_N\p{g_S}
    : \t{all elements of $k$ are distinct} },
  \label{eq:index_values}
\end{align}
where each $k\in\C_N\p{g}$ has the natural decomposition
$k=\p{k_S:S\in\PP^*\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$.  In words,
$\C_N\p{g}$ consists of all the ways to choose, for each subset $S$ of
the tensors $\set{w_j}$, a set $k_S$ of $g_S$ distinct spins, with the
restriction that all $k_S$ are orthogonal: if $z\in k_S$, then
$z\notin k_R$ for $R\ne S$.  Having defined the set $\C_N\p{g}$ of
index values, we define the restriction of $k\in\C_N\p{g}$ to values
associated with $w_j$ and $X_j$:
\begin{align}
  k_j \equiv \bigcup_{S\in\PP^*\p{\ZZ_p} : j\in S} k_S
  = \bigcup_{R\in\PP^*\p{\ZZ_p\setminus\set{j}}} k_{\set{j}\cup R},
\end{align}
where the union of lists $k_S$ denotes a concatenation of those lists.
These definitions allow us to expand
\begin{align}
  w\p{g} \equiv \sum_{k\in\C_N\p{g}} \prod_{j\in\ZZ_p} w_j\p{k_j},
  &&
  X\p{g} \equiv \prod_{j\in\ZZ_p}
  X_j\p{\ell_j}~\t{for some}~\ell\in\C_N\p{g},
  \label{eq:diagram_factors}
\end{align}
where the choice of $\ell\in\C_N\p{g}$ does not matter after
projection onto permutationally symmetric manifold.  Finally, we can
work out that
\begin{align}
  \S\p{g} \equiv \sp{\prod_{j\in\ZZ_p} \C\p{ g_S : j\in S }}
  \sp{\prod_{S\in\PP^*\p{\ZZ_p}} \p{g_S!}^\abs{S}}
  = \prod_{j\in\ZZ_p} M_j!,
  \label{eq:sym_prod_symmetry}
\end{align}
where the first product in \eqref{eq:sym_prod_symmetry} contains
multinomial coefficients
\begin{align}
  \C\p{a_1,a_2,\cdots,a_m}
  \equiv \f{\p{a_1+a_2+\cdots+a_m}!}{a_1!a_2!\cdots a_m!}
\end{align}
that account for the number of ways to partition the indices of each
$w_j$ into sets of shared indices as specified by $g$, and the second
product in \eqref{eq:sym_prod_symmetry} accounts for the number of
ways to permute each set of shared indices on every tensor.
Altogether, the symmetry factors $\S\p{g}$ cancel out with the
$\sim M_j!$ prefactors in \eqref{eq:sym_prod_vec}, so
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} X_j\p{k}
  \EQPS \sum_{g\in\G\p{\m M}} w\p{g} X\p{g}.
  \label{eq:sym_prod_group}
\end{align}
This result essentially gives us a diagrammatic expansion for the
product of multi-body operators, projected onto the permutationally
symmetric manifold.

>>>>

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\I}{\mathcal{I}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\Z}{\mathcal{Z}}

@@

In order to induce non-trivial dynamics for initially PS states
$\ket\psi\in\M_0$, which are otherwise stationary eigenstates of the
interaction Hamiltonian $H_0$ in \eqref{eq:H_0}, we will add to the
Hamiltonian an operator of the general form \red{[TODO: consider more
  general sums of such operators]}
\begin{align}
  X_{\m w} \equiv \sum_{k\in\C_N\p{M}} w_k X_k,
  \label{eq:pert}
\end{align}
where $X$ is an $M$-spin operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
an ordered choice of $M$ spins that $X_k$ acts on; $\C_N\p{M}$ is the
set of all such choices of $M$ elements from a set of size $N$; and
$\m w$ is a dimension-$M$ tensor with scalar entries $w_k$.  Even
though only ${N\choose M}$ entries of $\m w$ are used to define
$X_{\m w}$ in \eqref{eq:pert}, for later convenience we define $w_k$
for all indices $k\in\ZZ_N^M$ by enforcing that $\m w$ is symmetric
with respect to any permutation of its indices, and that $w_k=0$ for
any $k$ with repeated values.  In contrast, $X_k$ generally need not
be symmetric with respect to permutations of $k$, although we will
still define $X_k=0$ for any $k$ with repeated values.  A chiral spin
model, for example, may have two-body interactions of the form
$X_{pq}=s_\x^{(p)}s_\y^{(q)}-s_\y^{(p)}s_\x^{(q)}$, which are not
invariant under exchange of spins $p$ and $q$.

@@

Defining $\m w_\Delta$ to be the projection of the tensor $\m w$ onto
the $\Delta$-eigenspace of $\hat{\m h}$, we can expand
\begin{align}
  \m w = \sum_\Delta \m w_\Delta,
  &&
  \hat{\m h}\p{\m w} = \sum_\Delta \Delta \m w_\Delta,
\end{align}
and in turn
\begin{align}
  X_{\m w} = \sum_\Delta X_{\m w_\Delta},
  &&
  X_{\hat{\m h}\p{\m w}} = \sum_\Delta \Delta X_{\m w_\Delta}.
\end{align}
Substituting these expansions into \eqref{eq:diagnosis}, we find
\begin{align}
  H_0 X_{\m w} \ket\psi
  = \sum_\Delta\p{E_0 + \Delta} X_{\m w_\Delta} \ket\psi,
\end{align}
which implies that the action of $X_{\m w_\Delta}$ on a PS state
$\ket\psi\in\M_0$ generates a state of definite interaction energy
$E_0+\Delta$.  It follows that
$\P_\Delta X_{\m w} \P_0 = \P_\Delta X_{\m w_\Delta} \P_0 = X_{\m
  w_\Delta} \P_0$, and so
\begin{align}
  H_{\t{eff}}^{(2)} = -\sum_{\Delta\ne0}
  \f1\Delta \P_0 X_{\m w_\Delta} X_{\m w_\Delta} \P_0.
\end{align}
The product $\P_0 X_{\m w_\Delta} X_{\m w_\Delta} \P_0$ generally
includes $r$-body operators for all positive integers $r\le2M$.  We
provide a general prescription for simplifying such products in
Appendix \ref{sec:operator_product}, and explicitly work out the
single-body ($M=1$) and two-body ($M=2$) cases in Appendices
\ref{sec:single_pair_prod} and \ref{sec:two_pair_prod}.  If $X$ is a
single-body operator ($M=1$), then
\begin{align}
  H_{\t{eff}}^{(2)}
  = \sum_{\Delta\ne0} \f{\var\p{\m w_\Delta}}{\p{N-1}\Delta}
  \p{\col{X}^2 - N \col{X^2}},
  &&
  \var\p{\m m} \equiv \f1N \sum_{p\in\ZZ_N} \p{m_p-\bar m}^2,
\end{align}
which in the SU(2) case implies that an inhomogeneous magnetic field
($X\propto\sigma_\z$) yields an OAT Hamiltonian at second order
perturbation theory.

@@

In order to project an operator $X_{\m w}$ onto
$\ul{\M}_{r_{\t{max}}}$, we first need to construct a basis for
$\ul{\M}_{r_{\t{max}}}$, and then compute all matrix elements of
$X_{\m w}$ in the constructed basis.  Our strategy will be to first
identify a basis for $\M_0$, and then find, for each
$r=1,\cdots,r_{\t{max}}$, a suitable set of $r$-body operators
$\set{O}$ and dimension-$r$ tensors $\set{\m u}$ for which the states
$\ket{O,\m u,m}\propto O_{\m u}\ket{m}$ form an orthonormal basis for
$\M_r$.  Matrix elements of $X_{\m w}$ with respect to this basis for
$\ul{\M}_{r_{\t{max}}}$ then take the form
\begin{align}
  \bk{O,\m u,\ell|X_{\m w}|Q,\m v,m}
  = \f{\bk{\ell|O_{\m u}^\dag X_{\m w} Q_{\m v}|m}}
  {\sqrt{\bk{\ell|O_{\m u}^\dag O_{\m u}|\ell}
      \bk{m|Q_{\m v}^\dag Q_{\m v}|m}}}.
  \label{eq:matrix_element}
\end{align}
In order to compute these matrix elements, we simplify products of the
form $\P_0 O_{\m u}^\dag X_{\m w} Q_{\m v} \P_0$ and
$\P_0 O_{\m u}^\dag O_{\m u} \P_0$, with $\P_0$ a projector onto
$\M_0$, in Appendix \ref{sec:operator_product}.

@@

\vspace{3cm}

% For consistency with \eqref{eq:matrix_element}, for each state
% $\ket{m}$ in this basis for $\M_0$ we define
% $\ket{\1,1,m}\equiv\ket{m}$, with $\1$ the identity operator, and we
% define $\1_1\equiv\1$.

@@

In this section, we simplify the products of the form
$H_0X_{\m w}\ket\psi$ to arrive at the expansion in
\eqref{eq:diagnosis}, which defines conditions under which a
multi-body perturbation $X_{\m w_\Delta}$ generates eigenstates of the
SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0} when
acting on a permutationally symmetric state $\ket\psi\in\M_0$.  Here
$X_{\m w}$ is an $M$-body perturbation defined by
\begin{align}
  X_{\m w} \equiv \sum_{k\in\C_N\p{M}} w_k X_k,
  \label{eq:pert_copy}
\end{align}
where $X$ is an $M$-spin operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
an ordered choice of $M$ spins that $X_k$ acts on; $\C_N\p{M}$ is the
set of all such choices of $M$ elements from a set of size $N$; and
$\m w$ is a dimension-$M$ (i.e.~$M$-index) tensor with scalar entries
$w_k$.  Even though only ${N\choose M}$ entries of $\m w$ are used to
define $X_{\m w}$ in \eqref{eq:pert_copy}, for later convenience we
define $w_k$ for all indices $k\in\ZZ_N^M$ by enforcing that $\m w$ is
symmetric with respect to any permutation of its indices, and that
$w_k=0$ for any $k$ with repeated values.  In contrast, $X_k$
generally need not be symmetric with respect to permutations of $k$,
although we will still define $X_k=0$ for any $k$ with repeated
values.

@@

\red{[TODO: fix derivation below to account for operators $X$ that are
  not permutationally symmetric]}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Permutation-symmetrized products of multi-body operators}
\label{sec:operator_product}

@@

Given a system of $N$ spins, we wish to project a product of
multi-body operators onto the permutationally symmetric manifold (PS)
$\M_0$.  For $p$ multi-body operators, such a product takes the form
\begin{align}
  \m X_{\m w}
  \equiv \prod_{\alpha\in\ZZ_p} \sum_{k_\alpha\in\C_N\p{M_\alpha}}
  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
  \label{eq:sym_prod_start}
\end{align}
where $\m X\equiv\p{X_1,X_2,\cdots,X_p}$ is a vector of
$M_\alpha$-spin operators $X_\alpha$;
$\m w\equiv\p{w_1,w_2,\cdots,w_p}$ is a vector of dimension-$M_\alpha$
tensors $w_\alpha$;
$k_\alpha\equiv\p{k_{\alpha,1},k_{\alpha,2},\cdots,k_{\alpha,M_\alpha}}$
is an ordered choice of $M_\alpha$ spins that $X_\alpha\p{k_\alpha}$
acts on; $\C_N\p{M_\alpha}$ is the set of all such choices of
$M_\alpha$ elements from a set of size $N$; and $w_\alpha\p{k_\alpha}$
is a scalar entry of $w_\alpha$.  Although only ${N \choose M_\alpha}$
entries of $w_\alpha$ are used in \eqref{eq:sym_prod_start}, we define
$w_\alpha\p{k_\alpha}$ for all $k_\alpha\in\ZZ_N^{M_\alpha}$ by
enforcing that $w_\alpha\p{k_\alpha}$ is symmetric with respect to
permutations of $k_\alpha$, and that $w_\alpha\p{k_\alpha}=0$ if
$k_\alpha$ has any repeated values.  In contrast,
$X_\alpha\p{k_\alpha}$ generally need not be symmetric with respect to
permutations of $k_\alpha$, although we still enforce that
$X_\alpha\p{k_\alpha}=0$ if $k_\alpha$ has any repeated values.

To simplify the product in \eqref{eq:sym_prod_start}, we rewrite the
product of sums as a sum of products, and symmetrize over all
permutations of $k_\alpha\in\C_N\p{M_\alpha}$ in order to make all
individual sums independent:
\begin{align}
  \m X_{\m w}
  &= \f1{M_1!} \sum_{k_1\in\ZZ_N^{M_1}}
  \f1{M_2!} \sum_{k_2\in\ZZ_N^{M_2}} \cdots
  \f1{M_p!} \sum_{k_p\in\ZZ_N^{M_p}}
  \prod_{\alpha\in\ZZ_p} w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha} \\
  &= \f1{\m M!} \sum_{k\in\ZZ_N^{\m M}} \prod_{\alpha\in\ZZ_p}
  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
  \label{eq:sum_of_products}
\end{align}
where $\m M\equiv\p{M_1,M_2,\cdots,M_p}$ is a vector of the
dimensions $M_\alpha$; $\m M!\equiv\prod_{M_\alpha\in\m M}M_\alpha!$
for shorthand; and
$\ZZ_N^{\m M}\equiv\bigotimes_{\alpha\in\ZZ_p}\ZZ_N^{M_\alpha}$, such
that each $k\in\ZZ_N^{\m M}$ has the decomposition
$k=\p{k_1,k_2,\cdots,k_p}$ with $k_\alpha\in\ZZ_N^{M_\alpha}$.

We now classify terms in \eqref{eq:sum_of_products} by the numbers
$g_S$ of indices shared by all tensors $w_\alpha$ with
$\alpha\in S\subset\ZZ_p$.  For example, a term of the form
$w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct indices
$a,b,c,d,e$ would have
\begin{align}
  g_{\set{1}} = 1,
  &&
  g_{\set{1,2,3}} = 1,
  &&
  g_{\set{1,3}} = 1,
  &&
  g_{\set{2,3}} = 2,
\end{align}
and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
assignment can be associated with the Venn diagram
\begin{align}
  \diagram{example_123},
  \label{eq:venn_diagram}
\end{align}
where $g_S$ is determined the number of dots at the intersection of
circles $\alpha\in S$.  We denote the set of all nonempty subsets of
$\ZZ_p$ by $\PP^*\p{\ZZ_p}$, and denote a choice of $g_S$ for all
$S\in\PP^*\p{\ZZ_p}$ by $g$.  With representations such as
\eqref{eq:venn_diagram} in mind, we will loosely refer to $g$ as
simply a ``diagram''.  We define
$\abs{g}\equiv\sum_{S\in\PP^*\p{\ZZ_p}}g_S$ to be the number of
indices considered in $g$ (or equivalently the number of dots in the
diagram $g$), and we denote the set of all diagrams for
$p\equiv\dim\p{\m M}$ tensors with dimensions $\m M$ by $\G\p{\m M}$.
For consistency, all diagrams $g\in\G\p{\m M}$ must associate exactly
$M_\alpha$ indices with the dimension-$M_\alpha$ tensor $w_\alpha$,
i.e.
\begin{align}
  \sum_{S\in\PP^*\p{\ZZ_p}\,:\,\alpha\in S} g_S
  = \sum_{R\in\PP^*\p{\ZZ_p\setminus\set{\alpha}}} g_{\set{\alpha}\cup R}
  = M_\alpha
\end{align}
for all $\alpha\in\ZZ_p$.  Schematically, a classification of the
terms in \eqref{eq:sum_of_products} by diagrams $g\in\G\p{\m M}$ will
allow us to expand
\begin{align}
  \m X_{\m w} \EQPS \sum_{g\in\G\p{\m M}} \m w\p{g} \m X\p{g},
  \label{eq:sym_prod_teaser}
\end{align}
where $\EQPS$ denotes equality up to a restriction to the PS manifold
$\M_0$; $\m w\p{g}$ is a scalar acquired by assigning $\abs{g}$
indices to $\m w$ according to $g$, and summing over those indices in
some appropriate manner; and $\m X\p{g}$ is a $\abs{g}$-spin operator
acquired by averaging over all of the distinct ways to assign
$\abs{g}$ indices to $\m X$ according to $g$ (clarified below).

In order to derive the expansion in \eqref{eq:sym_prod_teaser} and
find the explicit forms of $\m w\p{g}$ and $\m X\p{g}$, we need to
collect all terms in \eqref{eq:sum_of_products} that are associated
with the diagram $g$.  To this end, we first consider all distinct
ways to assign $\abs{g}$ indices to specific tensor factors of $\m w$
and $\m X$ in a manner consistent with $g$.  For example, if
$\m X=\p{X_1,X_2}$ with dimensions $\p{M_1,M_2}=\p{2,1}$ and
$\abs{g}=2$ with $g_{\set{1,2}}=g_{\set{2}}=1$, then
$X_1\p{a,b} X_2\p{b}$ and $X_1\p{b,a} X_2\p{b}$ are distinct
assignments of $\abs{g}$ indices to $\m X$ in a manner consistent with
$g$, provided that $a\ne b$.  After assigning $\abs{g}$ indices to
$\m w$ and $\m X$ in a manner consistent with $g$, one needs to sum
over all values that these indices can take, which amounts to a sum
over all $k=\p{k_1,k_2,\cdots,k_{\abs{g}}}\in\ZZ_N^{\abs{g}}$ with the
restriction that $k_\mu\ne k_\nu$ for all $\mu\ne\nu$.

A specific assignment of indices can be thought of as a choice, for
each dot in a region $S$ of a diagram $g$, of which tensor factor of
$w_\alpha$ and $X_\alpha$ (for all $\alpha\in S$) that dot is
associated with.  The only catch is that such choices treat all dots
in $g$ as distinct, whereas the $g_S$ dots within region $S$ of $g$
are actually indistinguishable.  The assignment
$X_1\p{a,b,c} X_2\p{a,b}$, for example, is the same as
$X_1\p{b,a,c} X_2\p{b,a}$.  We denote the set of all distinct index
assignments according to $g$ by $\L\p{g}$.  The number of such
assignments is
  \abs{\L\p{g}}
  = \sp{\prod_{\alpha\in\ZZ_p} \C\p{ g_S : \alpha\in S }}
  \sp{\prod_{S\in\PP^*\p{\ZZ_p}} \p{g_S!}^{\abs{S}}}
  \sp{\prod_{S\in\PP^*\p{\ZZ_p}} g_S!}^{-1}
  = \m M! \sp{\prod_{S\in\PP^*\p{\ZZ_p}} g_S!}^{-1},
  \label{eq:index_assignment_num}
where the first product in \eqref{eq:index_assignment_num} contains
multinomial coefficients
  \C\p{a_1,a_2,\cdots,a_m}
  \equiv \f{\p{a_1+a_2+\cdots+a_m}!}{a_1!a_2!\cdots a_m!}
that account for the number of ways to partition the $M_\alpha$
indices of each $w_\alpha$ and $X_\alpha$ into sets of shared indices
as specified by $g$, the second product in
\eqref{eq:index_assignment_num} accounts for the number of ways to
permute each set of shared indices on each $w_\alpha$ and $X_\alpha$,
and the third product in \eqref{eq:index_assignment_num} accounts for
the treatment of the dots within each region $S$ of $g$ as
distinguishable, when in fact they are not.  Altogether, we have
  \m X_{\m w}
  = \f1{\m M!} \sum_{g\in\G\p{\m M}} \sum_{\ell\in\L\p{g}}
  \sum_{\substack{k\in\ZZ_N^{\abs{g}}\\
      k_\mu\ne k_\nu\forall \mu\ne\nu}}
  \m w\p{k,\ell} \m X\p{k,\ell},
where $\m w\p{k,\ell}$ and $\m X\p{k,\ell}$ denote the assignment of
indices $k$ to all $w_\alpha$ and $X_\alpha$ according to $\ell$.  The
invariance of each tensor $w_\alpha$ under arbitrary permutations of
its indices means that all assignments of $\abs{g}$ indices to $\m w$
consistent with $g$ are equivalent,
e.g.~$w_1\p{a,b} w_2\p{b} = w_1\p{b,a} w_2\p{b}$.  The value of
$\m w\p{k,\ell}$ therefore depends on $g$, but not the specific choice
of $\ell\in\L\p{g}$.  Furthermore, the specific choice of spins $k$ in
$\m X\p{k,\ell}$ does not matter after projection onto the PS manifold
$\M_0$, so the dependence of $\m X\p{k,\ell}$ on $k$ is similarly
superfluous, which is to say that we can write
  \m X_{\m w}
  \EQPS \f1{\m M!} \sum_{g\in\G\p{\m M}}
  \sum_{\substack{k\in\ZZ_N^{\abs{g}}\\
      k_\mu\ne k_\nu\forall \mu\ne\nu}}
  \m w\p{k,g} \sum_{\ell\in\L\p{g}} \m X\p{\ell}.
Defining the average of $\m X\p{\ell}$ over all $\ell\in\L\p{g}$,
  \m X\p{g} \equiv \f1{\abs{\L\p{g}}}
  \sum_{\ell\in\L\p{g}} \m X\p{\ell},
we can more simply write
  \m X_{\m w}
  \EQPS \f1{\m M!} \sum_{g\in\G\p{\m M}} \abs{\L\p{g}}
  \sum_{\substack{k\in\ZZ_N^{\abs{g}}\\
      k_\mu\ne k_\nu\forall \mu\ne\nu}}
  \m w\p{k,g} \m X\p{g}.
We now examine the sum over all $k\in\ZZ_N^{\abs{g}}$ with
$k_\mu\ne k_\nu\forall \mu\ne\nu$.  For any given instance of
$\m w\p{k,g}$, each entry $k_\mu\in k$ is associated with tensors
$\set{w_\alpha:\alpha\in S}$ for some subset $S\in\PP^*\p{\ZZ_p}$.
Identifying the indices associated with $S\in\PP^*\p{\ZZ_p}$ by
$k_S\in\ZZ_N^{g_S}$ and enforcing the decomposition
$k=\p{k_S:S\in\PP^*\p{\ZZ_p}}$, we can write
